The Hyperedge Event Model

### 2018-10-31 paper/revision1/ba-sample.tex
added subsection discussing HEM in the context of existing models

	We develop a statistical dynamic network model, which we term the hyperedge event model (HEM), that integrates the two components that govern hyperedge event formation: (1) the formation of the vertices that are incident to the hyperedge, and (2) the timing of the hyperedge event. Though the HEM, based on our knowledge of the literature, is unique in modeling hyperedge events in continuous time, we build our model using features inspired by many of the existing models for dynamic network data. First, the features that govern which receivers are selected by a given sender are defined using subgraph configuration counts (e.g., the number of triangles formed through sending the tie, the number of previous hyperedges that would be reciprocated by the tie). The HEM's use of subgraph configuration formation follows very similar aspects of the continuous-time event model proposed by \cite{PerryWolfe2012}, [HERE]

In what follows, we define the HEM's generative process for hyperedge event data (Section \ref{sec:generative process}), derive the conditional posteriors for Bayesian inference, and present tests of our software implementation (Section \ref{sec:inference}). We then demonstrate our model's applicability in a case study (Section \ref{sec:Emails}) where we analyze a corpus of internal county government emails and illustrate how to perform model selection, posterior predictive checks, and exploratory analysis using our model. We conclude in Section \ref{sec:conclusion}.~\looseness=-1

@@

					\IF{$n \geq 2$ tied events} 
					\STATE	set $s_e,\ldots, s_{e+n-1}=\mbox{argmin}_{i}(\tau_{ie}),$
					\STATE	set $\boldsymbol{r}_e=\boldsymbol{u}_{s_e e},\ldots,\boldsymbol{r}_{e+n-1}=\boldsymbol{u}_{s_{e+n-1} e}$
					\STATE	set $t_e, \ldots, t_{e+n-1}=t_{e-1} + \min_i\tau_{ie}$
					\STATE		jump to $e = e+n$
					\ELSE
					\STATE	set $s_e= \mbox{argmin}_{i}(\tau_{ie})$
					\STATE		set $\boldsymbol{r}_e = \boldsymbol{u}_{s_e e}$
					\STATE	set $t_e =t_{e-1} + \min_i\tau_{ie}$
					\ENDIF
					\ENDFOR
				\end{algorithmic}
				\label{alg:generative}
			\end{algorithm}
	\begin{equation}
		\begin{aligned}
			s_e &= \mbox{argmin}_{i}(\tau_{ie}),\\
			\boldsymbol{r}_e &= \boldsymbol{u}_{s_e e},\\
			t_e &=t_{e-1} + \tau_{s_e e}.
		\end{aligned}
	\end{equation}
		Therefore, the HEM assumes a sender-driven process---i.e., the receivers and timestamp of an event are jointly determined by the sender's urgency to direct the event to those selected receivers---which is similar to \cite{snijders1996stochastic} in that each actor gets his/her own rate function to determine ``who will make the next change in network." Note that our generative process allows for tied events. In the case of tied events---i.e., multiple senders draw exactly the same candidate timestamps---we assume that all events are generated and occur simultaneously. Algorithm \ref{alg:generative} summarizes the entire generative process for hyperedge events with one sender and one or more receivers.~\looseness=-1

>>>>

	We develop a statistical dynamic network model, which we term the hyperedge event model (HEM), that integrates the two components that govern hyperedge event formation: (1) the formation of the vertices that are incident to the hyperedge, and (2) the timing of the hyperedge event. In what follows, we define the HEM's generative process for hyperedge event data (Section \ref{sec:generative process}), derive the conditional posteriors for Bayesian inference, and present tests of our software implementation (Section \ref{sec:inference}). We then demonstrate our model's applicability in a case study (Section \ref{sec:Emails}) where we analyze a corpus of internal county government emails and illustrate how to perform model selection, posterior predictive checks, and exploratory analysis using our model. We conclude in Section \ref{sec:conclusion}.~\looseness=-1

@@

					\IF{$n \geq 2$ tied events} 
					\STATE	set $s_e,\ldots, s_{e+n-1}=\mbox{argmin}_{i}(\tau_{ie}),$
					\STATE	set $\boldsymbol{r}_e=\boldsymbol{u}_{s_e e},\ldots,\boldsymbol{r}_{e+n-1}=\boldsymbol{u}_{s_{e+n-1} e}$
					\STATE	set $t_e, \ldots, t_{e+n-1}=t_{e-1} + \min_i\tau_{ie}$
					\STATE		jump to $e = e+n$
					\ELSE
					\STATE	set $s_e= \mbox{argmin}_{i}(\tau_{ie})$
					\STATE		set $\boldsymbol{r}_e = \boldsymbol{u}_{s_e e}$
					\STATE	set $t_e =t_{e-1} + \min_i\tau_{ie}$
					\ENDIF
					\ENDFOR
				\end{algorithmic}
				\label{alg:generative}
			\end{algorithm}
	\begin{equation}
		\begin{aligned}
			s_e &= \mbox{argmin}_{i}(\tau_{ie}),\\
			\boldsymbol{r}_e &= \boldsymbol{u}_{s_e e},\\
			t_e &=t_{e-1} + \tau_{s_e e}.
		\end{aligned}
	\end{equation}
		Therefore, the HEM assumes a sender-driven process---i.e., the receivers and timestamp of an event are jointly determined by the sender's urgency to direct the event to those selected receivers---which is similar to \cite{snijders1996stochastic} in that each actor gets his/her own rate function to determine ``who will make the next change in network." Note that our generative process allows for tied events. In the case of tied events---i.e., multiple senders draw exactly the same candidate timestamps---we assume that all events are generated and occur simultaneously. Algorithm \ref{alg:generative} summarizes the entire generative process for hyperedge events with one sender and one or more receivers.     ~\looseness=-1

\subsection{Basis in existing models}

Though the HEM, based on our knowledge of the literature, is unique in modeling hyperedge events in continuous time, we build our model using features inspired by many of the existing models for dynamic network data. First, the features that govern which receivers are selected by a given sender are defined using subgraph configuration counts (e.g., the number of triangles formed through sending the tie, the number of previous hyperedges that would be reciprocated by the tie). The HEM's use of subgraph configuration formation follows a similar process underlying network change in the the continuous-time event model proposed by \cite{PerryWolfe2012} as well as the relational event model \citep{Butts2008}, and is also related to the subgraph completion processes used to model discrete-time network change in the temporal ERGM \citep{hanneke2010discrete}. Second, the use of candidate receivers as a mechanism through which continuous network change is modeled is related to the use of latent changes between snapshot network observations in the stochastic actor oriented model (SAOM) by \cite{snijders1996stochastic}.  Third, also like the SAOM, the HEM includes separate sets of parameters that are used to model event timing and edge generation. Though unique in modeling hyperedges, the HEM is grounded in existing continuous-time models of networks, ERGM-style models, and the SAOM.

The models discussed above are those, like HEM,  in which edge formation is driven by subgraph structure. There is another class of network models, originating with the latent space model introduced by \cite{hoff2002}, in which tie formation is modeled by embedding nodes in a space of latent attributes. Under this modeling framework, each node is attributed with a coordinate in a metric space, and the likelihood of a tie between two nodes is inversely related to the distance between the two nodes in the metric space. The latent space model has seen several extensions, the latest of which tis the latent factor model proposed by \cite{minhas2016inferential}. A number of dynamic extensions of the latent space framework have been developed \citep[e.g., ][]{sarkar2005dynamic,sewell2015latent,ward2013gravity}. The HEM, as we have presented it, does not involve latent node attributes, but in future work the HEM could be extended to incorporate latent node features---resulting in a latent space model for dynamic hyperedges.
